Capstone Movielens Report

### 2025-06-02 reports/capstone-movielens.site/02.8-0.UMGYDE-model.Rmd
UMGYDE Model Analysis: complted.

#### UMGYDE Model Regularization: Support Functions
\

@@

::: { #func.regularize.test_lambda.UMGYD_effect.cv.details .sidebar }
##### Details
\
Under the hood, the function internally calls the [regularize.train_UMGYD_effect.cv](#func.regularize.train_UMGYD_effect.cv) and then [calc_UMGY_SmoothedDay_effect.RMSE.cv](#func.calc_UMGY_SmoothedDay_effect.RMSE.cv) described above, and returns the _RMSE_ value that the latter inner function calculates for the given [lambda](#func.regularize.test_lambda.UMGYD_effect.cv.params.lambda) parameter value passed to the former one.
:::

@@

@@

Below we provide the most significant part of the code that performs this operation:
```{r, eval=FALSE}
```
```{r}
str(cv.UMGYDE.preset.result)
cv.UMGYDE.preset.result$best_result
```

::: {.noteblock data-latex=""}
The complete version of the source code provided in this section can be found in the [UMGYDE Model Regularization: Pre-configuration](https://github.com/AzKurban-edX-DS/Capstone-MovieLens/blob/main/r/src/capstone-movielens.main.R#L1493) section of the [capstone-movielens.main.R](https://github.com/AzKurban-edX-DS/Capstone-MovieLens/blob/main/r/src/capstone-movielens.main.R#L1493) script on _GitHub_.
:::

@@

::: {.noteblock data-latex=""}
The complete version of the source code provided in this section can be found in the [Fine-tuning Step of the Regularization Method for the User+Movie Model](https://github.com/AzKurban-edX-DS/Capstone-MovieLens/blob/main/r/src/capstone-movielens.main.R#L867) section of the [capstone-movielens.main.R](https://github.com/AzKurban-edX-DS/Capstone-MovieLens/blob/main/r/src/capstone-movielens.main.R) script on _GitHub_.
:::

@@

Below we provide the most significant part of the code that performs this operation:
```{r, eval=FALSE}
```
```{r}
str(UMGYDE.rglr.fine_tune.results)
UMGYDE.rglr.fine_tune.results$best_result
```

@@

::: {.noteblock data-latex=""}
The complete version of the source code provided in this section are available in the [Re-train Regularized UMGYDE Model for the best `lambda`](https://github.com/AzKurban-edX-DS/Capstone-MovieLens/blob/main/r/src/capstone-movielens.main.R#L1589) section of the [capstone-movielens.main.R](https://github.com/AzKurban-edX-DS/Capstone-MovieLens/blob/main/r/src/capstone-movielens.main.R#L1589) script on _GitHub_.
:::

Now, we can calculate the _Regularized UMGYDE Effect_ by retraining our model on the entire `edx` dataset with the best value of the $\lambda$ parameter we just calculated, for the definitive _Root Mean Squared Error_ calculation and use in subsequent models.

```{r, eval=FALSE}

::: {.noteblock data-latex=""}
Here we use the helper function [train_UMGYD_effect](#func.train_UMGYD_effect) described above in the [UMGYDE Model Regularization: Support Functions] section.
:::

We calculate the _Root Mean Squared Error_ for the ultimately computed _UMGYDE Effect_ using the helper function [calc_UMGYD_effect_RMSE.cv](#func.calc_UMGYD_effect_RMSE.cv) described above in the [UMGYDE Model Regularization: Support Functions] section as follows:
```{r, eval=FALSE}
```

Finally, we add the definitive result for the current model to our _Result Table_:
```{r, eval=FALSE}
```
```{r}
```

::: {.noteblock data-latex=""}
Now, we have a better result than one for the previous model, but still insufficient to meet the Project Objective.
:::

@@

>>>>

#### UMGYDE Model Regularization: Helper Functions
\

@@

::: { #func.regularize.test_lambda.UMGYD_effect.cv.details .sidebar }
###### Details
\
Under the hood, the function internally calls the [regularize.train_UMGYD_effect.cv](#func.regularize.train_UMGYD_effect.cv) and then [calc_UMGY_SmoothedDay_effect.RMSE.cv](#func.calc_UMGY_SmoothedDay_effect.RMSE.cv) described above, and returns the _RMSE_ value that the latter inner function calculates for the given [lambda](#func.regularize.test_lambda.UMGYD_effect.cv.params.lambda) parameter value passed to the former one.
:::

@@

::: {.noteblock data-latex=""}
The complete version of the source code shown in this section is available in the [UMGYDE Model Regularization: Pre-configuration](https://github.com/AzKurban-edX-DS/Capstone-MovieLens/blob/main/r/src/capstone-movielens.main.R#L2322) section of the [capstone-movielens.main.R](https://github.com/AzKurban-edX-DS/Capstone-MovieLens/blob/main/r/src/capstone-movielens.main.R#L2322) script on _GitHub_.
:::

@@

Below we provide the most significant part of the code that performs this operation:
```{r, eval=FALSE}
  lambdas <- seq(0, 256, 16)
  cv.UMGYDE.preset.result <- 
    tune.model_param(lambdas, regularize.test_lambda.UMGYD_effect.cv)
```
```{r}
str(cv.UMGYDE.preset.result)
cv.UMGYDE.preset.result$best_result
```

@@

::: {.noteblock data-latex=""}
The complete version of the source code shown in this section can be found in the [UMGYDE Model Regularization: Fine-tuning](https://github.com/AzKurban-edX-DS/Capstone-MovieLens/blob/main/r/src/capstone-movielens.main.R#L2378) section of the [capstone-movielens.main.R](https://github.com/AzKurban-edX-DS/Capstone-MovieLens/blob/main/r/src/capstone-movielens.main.R#L2378) script on _GitHub_.
:::

@@

Below we provide the most significant part of the code that performs this operation:
```{r, eval=FALSE}
endpoints <- 
  get_fine_tune.param.endpoints(cv.UMGYDE.preset.result$tuned.result)

UMGYDE.loop_starter <- c(endpoints["start"], 
                        endpoints["end"], 
                        8)

cache.base_name <- "UMGYDE.rglr.fine-tuning"

UMGYDE.rglr.fine_tune.results <- 
  model.tune.param_range(UMGYDE.loop_starter,
                         UMGYDE.rglr.fine_tune.cache.path,
                         cache.base_name,
                         regularize.test_lambda.UMGYD_effect.cv)
```
```{r}
str(UMGYDE.rglr.fine_tune.results)
UMGYDE.rglr.fine_tune.results$best_result
```

@@

::: {.noteblock data-latex=""}
The complete version of the source code shown in this section is available in the [UMGYDE Regularized Model: Retraining with the best params](https://github.com/AzKurban-edX-DS/Capstone-MovieLens/blob/main/r/src/capstone-movielens.main.R#L2417) section of the [capstone-movielens.main.R](https://github.com/AzKurban-edX-DS/Capstone-MovieLens/blob/main/r/src/capstone-movielens.main.R#L2417) script on _GitHub_.
:::

Now, we can compute the _Regularized UMGYDE Effect_ by re-training our model on the entire `edx` dataset with the best value of the $\lambda$ parameter we have just figured out and use it in subsequent analysis.

The most significant part of the code for this operation is shown below: 
```{r, eval=FALSE}
  best_result <- UMGYDE.rglr.fine_tune.results$best_result
  UMGYDE.rglr.best_lambda <- best_result["param.best_value"]
  UMGYDE.rglr.best_RMSE <- best_result["best_RMSE"]

  put_log1("Re-training Regularized User+Movie+Genre+Year+(Smoothed)Day Effect Model for the best `lambda`: %1...",
           UMGYDE.rglr.best_lambda)

  rglr.UMGYD_effect <- edx |> 
    regularize.train_UMGYD_effect(UMGYDE.rglr.best_lambda)
```
```{r}
  str(rglr.UMGYD_effect)
```

::: {.noteblock data-latex=""}
Here we use the helper function [regularize.train_UMGYD_effect](#func.regularize.train_UMGYD_effect) described above in the [UMGYDE Model: Helper Functions] section.
:::

We calculate the _Root Mean Squared Error_ for the ultimately computed _Regularized UMGYDE Effect_ using the helper function [calc_UMGY_SmoothedDay_effect.RMSE.cv](#func.calc_UMGY_SmoothedDay_effect.RMSE.cv) described above in the [UMGYDE Model: Helper Functions] section as follows:
```{r, eval=FALSE}
  rglr.UMGYD_effect.RMSE <- calc_UMGY_SmoothedDay_effect.RMSE.cv(rglr.UMGYD_effect)
```

As always, we add the definitive result for the current model to our _Result Table_:
```{r, eval=FALSE}
RMSEs.ResultTibble.UMGYDE.rglr.tuned <- RMSEs.ResultTibble.UMGYDE.tuned |> 
  RMSEs.AddRow("Regularized UMGYDE Model", 
               rglr.UMGYD_effect.RMSE,
               comment = "The best tuned and regularized UMGYDE Model.")
```
```{r}
RMSE_kable(RMSEs.ResultTibble.UMGYDE.rglr.tuned)
```

::: {.noteblock data-latex=""}
Now, we have a better result than the ones for the previous models, but still insufficient to meet the Project Objective. Let's see what else we can do to achieve our ultimate goal. 
:::

@@

#### UMGYDE Model: Final Holdout Test (Preliminary Assesment)

::: {.noteblock data-latex=""}
The complete version of the source code shown in this section is available in the [UMGYDE Model: Final Holdout Test](https://github.com/AzKurban-edX-DS/Capstone-MovieLens/blob/main/r/src/capstone-movielens.main.R#L2498) section of the [capstone-movielens.main.R](https://github.com/AzKurban-edX-DS/Capstone-MovieLens/blob/main/r/src/capstone-movielens.main.R#L2498) script on _GitHub_.
:::

Let's see what we have achieved so far, using the `final_holdout_test` dataset for the first time.
Firstly, we compute predictions for our current model:
```{r, eval=FALSE}
final.UMGYDE.predicted <- final_holdout_test |>
  UMGY_SmoothedDay_effect.predict(rglr.UMGYD_effect)
```
```{r}
str(final.UMGYDE.predicted)
```
::: {.noteblock data-latex=""}
Here we use the helper function [UMGY_SmoothedDay_effect.predict](#func.UMGY_SmoothedDay_effect.predict) described above in the [UMGYDE Model: Helper Functions] section.
:::

Then we compute the _RMSE_ and add the result to our _Result Table_:
```{r, eval=FALSE}
final.UMGYDE.predicted.RMSE <- rmse2(final_holdout_test$rating,
                                     final.UMGYDE.predicted$predicted)

final.RMSEs.ResultTibble.UMGYDE.rglr.tuned <- RMSEs.ResultTibble.UMGYDE.rglr.tuned |> 
  RMSEs.AddRow("Best UMGYDE Model (Final Test)", 
               final.UMGYDE.predicted.RMSE,
               comment = "Final Holdout Test of the best tuned and regularized UMGYDE Model.")
```
```{r}
RMSE_kable(final.RMSEs.ResultTibble.UMGYDE.rglr.tuned)
```

::: {.noteblock data-latex=""}
Here we use the helper function [rmse2](https://github.com/AzKurban-edX-DS/Capstone-MovieLens/blob/main/r/src/support-functions/common-helper.functions.R#L28) to calculate the _RMSE_ that is defined in the  [(R)MSE-related functions](https://github.com/AzKurban-edX-DS/Capstone-MovieLens/blob/main/r/src/support-functions/common-helper.functions.R#L18) section of the [common-helper.functions.R](https://github.com/AzKurban-edX-DS/Capstone-MovieLens/blob/main/r/src/support-functions/common-helper.functions.R#L28) script on _GitHub_.
:::

::: {.noteblock data-latex=""}
As expected, we have not achieved a result that is good enough to meet our _Project Objective_.
:::

### 2025-06-02 reports/capstone-movielens.site/04.appendix.Rmd
UMGYDE Model Analysis: complted.

Plots a graph based on the dataset passed in the [data](#func.data.plot.left.n.params.data) parameter, using both points and a line.

@@

::: { #func.data.plot.left.n.params.data .sidebar }
  - **data:** Dataset to use for the plot;

:::
::: { #func.data.plot.left.n.params.title .sidebar }
  - **title:** Title of the plot;

:::
::: { #func.data.plot.left.n.params.xname .sidebar }
  - **xname:** The name of the dataset column used as the source of the $x$ variable for the plot;

:::
::: { #func.data.plot.left.n.params.yname .sidebar }
  - **yname:** The name of the dataset column used as the source of the $y$ variable for the plot;

:::
::: { #func.data.plot.left.n.params.xlabel .sidebar }
  - *xlabel = NULL*:** The $x$ axis label. If `NULL`, the value of the `xname` parameter is used for the label;

:::
::: { #func.data.plot.left.n.params.ylabel .sidebar }
  - **ylabel = NULL:** The $y$ axis label. If `NULL`, the value of the `yname` parameter is used for the label.;

:::
::: { #func.data.plot.left.n.params.line_col .sidebar }
  - **line_col = blue:** The line color;

@@

The function is a wrapper for the [data.plot](#func.data.plot) function described above and is a simple and convenient tool for more detailed visualizing the left part of the graph.

@@

The soure code of the [data.plot.left.n](https://github.com/AzKurban-edX-DS/Capstone-MovieLens/blob/main/r/src/support-functions/data.helper.functions.R#L493) function is shown below:
```{r eval=FALSE}
data.plot.left.n <- function(data, 
                             left.n = 0,
                             title, 
                             xname, 
                             yname, 
                             xlabel, 
                             ylabel,
                             line_col = "red",
                             normalize = FALSE) {
  x_col <- data[, xname] 
  y_col <- data[, yname]

@@

Plots a graph based on the dataset passed in the [data](#func.data.plot.left_detailed.params.data) parameter, using both points and a line.

@@

::: { #func.data.plot.left_detailed.params.data .sidebar }
  - **data:** Dataset to use for the plot;

:::
::: { #func.data.plot.left_detailed.params.title .sidebar }
  - **title:** Title of the plot;

:::
::: { #func.data.plot.left_detailed.params.xname .sidebar }
  - **xname:** The name of the dataset column used as the source of the $x$ variable for the plot;

:::
::: { #func.data.plot.left_detailed.params.yname .sidebar }
  - **yname:** The name of the dataset column used as the source of the $y$ variable for the plot;

:::
::: { #func.data.plot.left_detailed.params.xlabel .sidebar }
  - *xlabel = NULL*:** The $x$ axis label. If `NULL`, the value of the `xname` parameter is used for the label;

:::
::: { #func.data.plot.left_detailed.params.ylabel .sidebar }
  - **ylabel = NULL:** The $y$ axis label. If `NULL`, the value of the `yname` parameter is used for the label.;

:::
::: { #func.data.plot.left_detailed.params.line_col .sidebar }
  - **line_col = blue:** The line color;

@@

The function is a wrapper for the [data.plot.left.n](#func.data.plot.left.n) function described above and is a simple and convenient tool for more detailed visualizing the left part of the graph.

>>>>

Plots a graph based on the first $n$ rows (the $n$ is specified by the [left.n](#func.data.plot.left.n.params.left.n) parameter) of the dataset passed in the [data](#func.data.plot.left.n.params.data) parameter, using both points and a line.

::: {.noteblock data-latex=""}
The dataset passed in the [data](#func.data.plot.left.n.params.data) parameter is assumed to be sorted by the column used for the $x$ axis (specified by the column name passed in the [xname](#func.data.plot.left.n.params.xname) parameter).
:::

@@

::: { #func.data.plot.left.n.params.data .sidebar }
  - **data:** Dataset to use for the plot.

:::
::: { #func.data.plot.left.n.params.left.n .sidebar }
  - **left.n:** If a positive numeric value, specifies the number of the first rows of the dataset to plot. Otherwise, the entire dataset data is set to plot.

:::
::: { #func.data.plot.left.n.params.title .sidebar }
  - **title:** Title of the plot.

:::
::: { #func.data.plot.left.n.params.xname .sidebar }
  - **xname:** The name of the dataset column used as the source of the $x$ variable for the plot.

:::
::: { #func.data.plot.left.n.params.yname .sidebar }
  - **yname:** The name of the dataset column used as the source of the $y$ variable for the plot.

:::
::: { #func.data.plot.left.n.params.xlabel .sidebar }
  - *xlabel = NULL*:** The $x$ axis label. If `NULL`, the value of the `xname` parameter is used for the label.

:::
::: { #func.data.plot.left.n.params.ylabel .sidebar }
  - **ylabel = NULL:** The $y$ axis label. If `NULL`, the value of the `yname` parameter is used for the label.

:::
::: { #func.data.plot.left.n.params.line_col .sidebar }
  - **line_col = blue:** The line color.

@@

The function is a wrapper for the [data.plot](#func.data.plot) function described above and is a simple and convenient tool for more detailed visualization of the left part of the graph, initially built based on the entire data set passed in the [data](#func.data.plot.left.n.params.data) parameter.
:::

@@

The source code of the [data.plot.left.n](https://github.com/AzKurban-edX-DS/Capstone-MovieLens/blob/main/r/src/support-functions/data.helper.functions.R#L493) function is shown below:
```{r eval=FALSE}
data.plot.left.n <- function(data, 
                             left.n = 0,
                             title, 
                             xname, 
                             yname, 
                             xlabel, 
                             ylabel,
                             line_col = "red",
                             normalize = FALSE) {
  x_col <- data[, xname] 
  y_col <- data[, yname]

@@

Combines two graphs plotted by the internally called helper functions in a grid: the first is a standard graph plotted by the [data.plot](#func.data.plot) function described above, the second is a graph based on the first $n$ rows (the $n$ is specified by the [left.n](#func.data.plot.left.n.params.left.n) parameter) of the dataset plotted by the [data.plot.left.n](#func.data.plot.left.n) function also described above (The data for both graphs is passed in the [data]).(#func.data.plot.left.n.params.data) parameter).

::: {.noteblock data-latex=""}
The dataset passed in the [data](#func.data.plot.left.n.params.data) parameter is assumed to be sorted by the column used for the $x$ axis (specified by the column name passed in the [xname](#func.data.plot.left.n.params.xname) parameter).
:::

@@

::: { #func.data.plot.left_detailed.params.data .sidebar }
  - **data:** Dataset to use for the plot.

:::
::: { #func.data.plot.left.n.params.left.n .sidebar }
  - **left.n:** If a positive numeric value, specifies the number of the first rows of the dataset to plot. Otherwise, the entire dataset data is set to plot.

:::
::: { #func.data.plot.left_detailed.params.title .sidebar }
  - **title:** Title of the plot.

:::
::: { #func.data.plot.left_detailed.params.xname .sidebar }
  - **xname:** The name of the dataset column used as the source of the $x$ variable for the plot.

:::
::: { #func.data.plot.left_detailed.params.yname .sidebar }
  - **yname:** The name of the dataset column used as the source of the $y$ variable for the plot.

:::
::: { #func.data.plot.left_detailed.params.xlabel1 .sidebar }
  - *xlabel1 = NULL*:** The $x$ axis label for the first plot. If `NULL`, the value of the `xname` parameter is used for the label.

:::
::: { #func.data.plot.left_detailed.params.xlabel2 .sidebar }
  - *xlabel2 = NULL*:** The $x$ axis label for the second plot. If `NULL`, the value of the `xname` parameter is used for the label.

:::
::: { #func.data.plot.left_detailed.params.ylabel1 .sidebar }
  - **ylabel1 = NULL:** The $y$ axis label for the first plot. If `NULL`, the value of the `yname` parameter is used for the label.

:::
::: { #func.data.plot.left_detailed.params.ylabel2 .sidebar }
  - **ylabel2 = NULL:** The $y$ axis label for the second plot. If `NULL`, the value of the `yname` parameter is used for the label.

:::
::: { #func.data.plot.left_detailed.params.line_col .sidebar }
  - **line_col = blue:** The line color.

@@

Under the hood, the function calls internally the [data.plot](#func.data.plot) and [data.plot.left.n](#func.data.plot.left.n) functions (described above) to build two graphs and combines them into a grid.
:::
